---
title: "Field_data"
author: "Ben Weinstein"
date: "4/20/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

The goal of this vignette is download data for individual trees in the sampling plots.

```{r}
library(ggplot2)
library(dplyr)
library(neonVegWrangleR)
```

NEON page: https://data.neonscience.org/data-products/DP1.10098.001

# Download NEON data and get UTM locations

```{r}
# load NEON woody vegetation structure data product into R
vst <- neonUtilities::loadByProduct("DP1.10098.001", check.size=F)

# calculate UTM coordinates of vst entries based on azimuth and distance
# measurements from plot reference points
vst_locations <- calc_tree_geolocations(vst, dataProd = "vst_mappingandtagging")
names(vst_locations)
```

Only tree growth forms

```{r}
trees<-vst_locations[["vst_apparentindividual"]] %>% filter(growthForm %in% c("single bole tree","multi-bole tree","small tree","sapling"))
```

# Clean location data to create unique ids

From NEON "Duplicates in vst_mappingandtagging may exist at the individualID level if errors have been corrected after ingest of the original record" 

How many duplicates locations do we have?
  
```{r}
head(vst_locations[["vst_mappingandtagging"]])
duplicateIDs<-vst_locations[["vst_mappingandtagging"]] %>% group_by(individualID) %>% summarize(n=n()) %>% filter(n>1)
dim(duplicateIDs)
```

NEON's recommended merge strategy is to use the more recent. For example:

```{r}
vst_locations[["vst_mappingandtagging"]] %>% filter(individualID %in% duplicateIDs$individualID[1])
```

```{r}
uniqueIDS<-vst_locations[["vst_mappingandtagging"]] %>% group_by(individualID) %>% arrange(desc(date)) %>% slice(1)
dim(uniqueIDS)
dim(vst_locations[["vst_mappingandtagging"]])
```

# Clean attribute table to get unique individuals per year

```{r}
paste("Original number of attribute records", nrow(trees))

duplicated_attributes<-trees %>% group_by(individualID,eventID) %>% summarize(n=n()) %>% filter(n>1)

duplicated_attributes %>% ungroup() %>% sample_n(size=4)
dim(duplicated_attributes)
```

### Show example

```{r}
vst_locations[["vst_apparentindividual"]] %>% filter(individualID=="NEON.PLA.D01.BART.00105",eventID=="vst_BART_2018")

vst_locations[["vst_apparentindividual"]] %>% filter(individualID=="NEON.PLA.D15.ONAQ.01775",eventID=="vst_ONAQ_2017")

#how different are the heights and dbh on average
height_diff<-trees %>% filter(individualID %in% duplicated_attributes$individualID) %>% group_by(individualID) %>% summarize(d=sum(diff(height)))
dim(height_diff)
ggplot(height_diff,aes(x=d)) + geom_density(fill="black") + labs(x="Sum difference in height among duplicated IDs for the same year (m)")
```

### Merge stratey

It is unknown the correct strategy here since many of these records are unsolvable. For the moment I am choosing to delete them.

```{r}
trees<-trees %>% filter(!individualID %in% duplicated_attributes$individualID)
```

# Merge Unique Locations and samples

From NEON:

"The vst_apparentindividual table contains one record per individualID per eventID, and includes growth form, structure and status data that may be linked to vst_mappingandtagging records via individualID"

```{r}
paste("Original number of attribute records", nrow(trees))

joined_df<-uniqueIDS %>% dplyr::select(c("individualID",colnames(uniqueIDS)[!colnames(uniqueIDS) %in% colnames(trees)])) %>% inner_join(trees,by="individualID")

paste("Joined number of attribute records", nrow(joined_df))
colnames(trees)[colnames(trees) %in% colnames(uniqueIDS)]
colnames(uniqueIDS)
head(joined_df)
```

```{r}
table(is.na(joined_df$height))
ggplot(joined_df) + geom_histogram(aes(x=height))
```


